---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Data wrangling

Load library

```{r}
library(here)
library(tidyverse)
```

Load data

```{r}
counts_row <- read_delim("data/counts-raw.txt.gz", delim = "\t")
```

Goals

subset rows of a dataframe and select columns of a dataframe

```{r}
research <- counts_row %>% filter(articleType == "Research Article")
```

fitler research articles from 2006

```{r}
research_2006 <- counts_row %>% filter(year == 2006) 
nrow(research_2006)
```

```{r}
research_2006_tweet <- research_2006 %>% filter(backtweetsCount == 0)
```

Reasearch articles from 2006 with a facebook comment
```{r}
research_2006_fb <- research_2006 %>% filter(year == 2006, facebookCommentCount > 0)
```

We can use logical operators:

AND & OR |

```{r}
research_2006_fb_tweet <- research %>% filter(year == 2006, facebookCommentCount > 0 | backtweetsCount > 0)
```

Article in 2006 with social
```{r}
research %>% filter(year == 2006, facebookCommentCount > 0 | backtweetsCount > 0, grepl("Infectious Diseases", plosSubjectTags))
```
Subset the columns with the article info using the dplyr function select()

```{r}
article_info <- research %>% select(doi, pubDate, journal, title, articleType, authorsCount)
```

New data frame with columns doi:authorsCount

```{r}
article_info <- research %>% select(doi:authorsCount)
```

Create a metrics dataframe

```{r}
metrics <- research %>% select(contains("Count"))
```

Exclude a column after contains

```{r}
metrics <- research %>% select(contains("Count"), -authorsCount)
metrics <- research %>% select(contains("Count"), -authorsCount, f1000Factor, wikipediaCites)
```

Journal column

```{r}
select(research, journal)
```

Slice: Slicing only the rows we want by numnbers

Rows 1:3 article info dataframe

```{r}
slice(article_info, 1:3)
```


```{r}
research_2009 <- counts_row %>% filter(year == 2009) %>% nrow()
research_2009 <- counts_row %>% filter(year == 2009) %>% 
    filter(backtweetsCount > 0 | facebookCommentCount > 0) %>% 
    nrow()

research_2009 <- counts_row %>% filter(year == 2009) %>% 
    filter(mendeleyReadersCount > 0) %>% 
    nrow()
    
```

```{r}
research_2009 <- counts_row %>% filter(year == 2009) %>% 
    filter(!is.na(backtweetsCount) & (backtweetsCount > 0))
```


Arrange to sort dataframe
```{r}
research %>% arrange(authorsCount, wosCountThru2011) %>% 
    select(authorsCount, wosCountThru2011) %>% 
    slice(1:10)
```

Decreasing order

```{r}
research %>% arrange(desc(authorsCount), desc(wosCountThru2011)) %>% 
    select(authorsCount, wosCountThru2011) %>% 
    slice(1:10)
```

```{r}
research %>% arrange(desc(wosCountThru2011)) %>% 
    select(wosCountThru2011, title) %>% 
    slice(1:3)


```

```{r}
research %>% arrange(desc(authorsCount)) %>% 
    select(authorsCount, title, journal, plosSubjectTags) %>% 
    slice(1:3)
```


## summarize data with dplyr

Goals:

* new columns with mutate
* summarize with summarize
* count number of obsevation n()
* group data by variable (group_by)

```{r}
research <- research %>% mutate(weeksSincePublished = daysSincePublished / 7)
```

```{r}
research <- research %>% mutate(weeksSincePublished = daysSincePublished / 7,
                                yearsSincePublised = weeksSincePublished / 52)
```

Summarize the data:

```{r}
research %>% summarise(plos_mean = mean(plosCommentCount))
```


```{r}
plot_stats <- research %>% summarise(plos_mean = mean(plosCommentCount), plos_sd <- sd(plosCommentCount))
```

```{r}
research %>% filter(journal == "pone", year == 2007) %>% 
    summarise(plos_mean = mean(plosCommentCount),
              plos_sd = sd(plosCommentCount))
```

add # of observations
```{r}
research %>% filter(journal == "pone", year == 2007) %>% 
    summarise(plos_mean = mean(plosCommentCount),
              plos_sd = sd(plosCommentCount),
              num = n())
```

Use group_by to get stats by a grouping of the data

```{r}
mean_metric_per_var <- function(metric, variable){
    result = integer(length(variable))
    for(v in variable){
        result[v] <- mean(metric(variable = v))
    }
    return(result)
}

research %>% group_by(journal) %>% 
    summarise(tweets_mean = mean(backtweetsCount))
```

Group by 2 variables

```{r}
research %>% group_by(journal, year) %>% 
    summarise(tweets_mean = mean(backtweetsCount))
```


```{r}
research %>% group_by(journal) %>% 
    summarise(num = n(),
              tweet_mean = mean(backtweetsCount),
        tweets_sem = sd(backtweetsCount)/sqrt(num))

```

Goals:

* Map columns of a data frame to plot aesthetics using aes
* Add layers to plot

```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount, y = wosCountThru2011)) +
    geom_point()

p <- ggplot(data = research, aes(x = pdfDownloadsCount, y = wosCountThru2011)) +
    geom_point(aes(color = journal))

```


Change size of points by variable

```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount, y = wosCountThru2011)) +
    geom_point(aes(size = wosCountThru2011))
```

Change alpha of points by variable

```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount, y = wosCountThru2011)) +
    geom_point(aes(alpha = daysSincePublished))
p

```


```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount, y = wosCountThru2011)) +
    geom_point(color = "red", aes(size = authorsCount))
p

```

Add another geometry layer

```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount, y = wosCountThru2011)) +
    geom_point(aes(alpha = daysSincePublished)) +
    geom_smooth()
p

```

Color both layers  by journal

```{r}
p <- ggplot(data = research, aes(x = pdfDownloadsCount, 
                                 y = wosCountThru2011,
                                 color = journal)) +
    geom_point() +
    geom_smooth() 
p
```


```{r}
p <- ggplot(data = research, aes(x = daysSincePublished,
                                 y = wosCountThru2011,
                                 color = journal)) +
    geom_point(aes(alpha = 0.5)) +
    geom_smooth(color = "red") 
p
```


```{r}
p + scale_x_log10() + scale_y_log10()
```

```{r}
p <- ggplot(data = research, aes(x = log10(daysSincePublished + 1),
                                 y = log10(wosCountThru2011 + 1))) +
    geom_point(aes(color = journal)) +
    geom_smooth() +
    scale_x_continuous(breaks = c(1,3), labels = c(10, 1000)) +
    scale_y_continuous(breaks = c(1,3), labels = c(10, 1000)) +
    scale_color_grey()
p

library(RColorBrewer)
p + scale_color_manual(volumes = )

p <- ggplot(data = research, aes(x = sqrt(pdfDownloadsCount),
                                 y = sqrt(wosCountThru2011))) +
    geom_point(aes(color = journal)) +
    geom_smooth(color = "red") +
    # scale_x_continuous(breaks = c(1,3), labels = c(10, 1000)) +
    # scale_y_continuous(breaks = c(1,3), labels = c(10, 1000)) +
    scale_color_brewer(palette = "Accent") +
    theme_bw(color = scale_color_brewer(palette = "Accent")) +
    labs(x = "Square root of downloads",
         y = "Square root of Counts through 2011")
p
```


